---
title: "Survival lecture 4"
author: "Steve Simon"
date: "April 8, 2018"
output: powerpoint_presentation
---

### Abstract

Lecture 4. Model fitting and diagnostics for the Cox model. In this lecture, you will work with more complex forms of the Cox model with multiple predictor variables. You'll include covariates in the Cox model to produce risk adjusted survival curves. You will also assess the underlying assumptions of the Cox model, particularly the assumption of proportional hazards.

<div class="notes">

This program was written on 2018-04-08 and last modified on `r Sys.Date()`.

This PowerPoint presentation was developed using R Markdown. The file does not need any special libraries other than the ones listed below. Many of the data sets in this program use data from Hosmer, Lemeshow, and May. I made one minor change, however, which was to force all the variable names to lower case.

</div>

### Peek at the first few rows of whas500

```{r load-libraries, echo=FALSE}
suppressMessages(suppressWarnings(library(broom)))
suppressMessages(suppressWarnings(library(dplyr)))
suppressMessages(suppressWarnings(library(ggplot2)))
suppressMessages(suppressWarnings(library(knitr)))
suppressMessages(suppressWarnings(library(magrittr)))
suppressMessages(suppressWarnings(library(survival)))
suppressMessages(suppressWarnings(library(tidyr)))
knitr::opts_chunk$set(
  echo=FALSE, 
  comment="",
  fig.width=4.5, 
  fig.height=2.5)
options(width=45)
```

```{r read}
fn <- "../data/whas500.RData" 
load(fn)
whas500$time_yrs <- whas500$lenfol/365.25
head(whas500, n=2)
```

<div class="notes">

We will be using the WHAS500 dataset. Here are the first couple of rows.

</div>

### Univariate models

```{r one-at-a-time}
coxph(Surv(time_yrs, fstat=="Dead")~age, data=whas500) %>%
  tidy %>%
  data.frame %>%
  kable
```

<div class="notes">

You should examine your independent variables one at a time before putting them all in a multivariate model.

</div>

```{r save-everything}
save.image("../data/class4.RData")
```